---
title: "Extreme Precip Analysis or a Grid over a Raster Grid Original Product"
output: html_notebook
---



Processing for Max Precip Daily Extremes Using LOCA Model Gridded Output.

Temperature is tricky since we will be using a Generalized Pareto to target the extreme values and that also requires a threshold to which the method is sensitive.  Temperature complicates this since we are looking at a goodly amount of non-stationarity as we move into the future.

Warning Typos are Legion.



# Load Required Libraries

```{r}
  library(package = "tidyverse")
  library(package = "tidypredict")

  library(package = "ncdf4")


  library(package = "lubridate") # processing dates and time
  library(package = "stringr")


  library(package = "reshape2")  # manipulating data frames
  library(package = "abind")

  library(package = "foreach")
  library(package = "parallel")


  
  
```


```{r}
detectCores(all.tests = FALSE, logical = TRUE)

```

# Available Ensembles

**Table for Available LOCA CMIP5 Point Stations**

CMIP5 Models        | CMIP5 Models          | CMIP5 Models
--------------------|-----------------------|---------------------
ACCESS1.0_r1i1p1    | ACCESS1.3_r1i1p1      | CCSM4_r6i1p1         
CESM1.BGC_r1i1p1    | CESM1.CAM5_r1i1p1     | CMCC.CMS_r1i1p1      
CMCC.CM_r1i1p1      | CNRM.CM5_r1i1p1       | CSIRO.Mk3.6.0_r1i1p1 
CanESM2_r1i1p1      | FGOALS.g2_r1i1p1      | GFDL.CM3_r1i1p1      
GFDL.ESM2G_r1i1p1   | GFDL.ESM2M_r1i1p1     | HadGEM2.AO_r1i1p1    
HadGEM2.CC_r1i1p1   | HadGEM2.ES_r1i1p1     | IPSL.CM5A.LR_r1i1p1  
IPSL.CM5A.MR_r1i1p1 | MIROC.ESM.CHEM_r1i1p1 | MIROC.ESM_r1i1p1     
MIROC5_r1i1p1       | MPI.ESM.LR_r1i1p1     | MPI.ESM.MR_r1i1p1    
MRI.CGCM3_r1i1p1    | NorESM1.M_r1i1p1      | bcc.csm1.1.m_r1i1p1


And we should include our time periods.  Here, we have a baseline period and a future period

# Data Loading and Pre-Processing

## File Control

```{r}

  URL_Root_Directory = "http://kyrill.ias.sdsmt.edu:8080/thredds/dodsC/LOCA_NGP/climatology/"
  URL_Root_Directory = "/maelstrom2/LOCA_GRIDDED_ENSEMBLES/LOCA_NGP/climatology/"
  
  # get URL files  "return_statistics_NGP_LOCA_rcp45_pr_2035-2064.RData""
  LOCA_File_Base  = 
  

```


## Data Ingest


```{r}

# Base Period 

  print("Importing Historical Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Base,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
    
    
#   Pull Coordinates
    
    Return_Intervals = return_intervals
    Ensemble         = ensemble_site
    Latitude         = latitude
    Longitude        = longitude
    
    All_Return_Intervals_xyesr_arr = array(data    = NA,
                                           dim     = c(length(Longitude),
                                                       length(Latitude),
                                                       length(Ensemble),
                                                       length(Scenario3),
                                                       length(Return_Intervals)),
                                           dimnames = list("Longitude"        = Longitude,
                                                           "Latitude"         = Latitude,
                                                           "Ensemble"         = Ensemble,
                                                           "Scenario"         = Scenario3,
                                                           "Return_Intervals" = Return_Intervals))
    
  
    
    All_Return_Intervals_xyesr_arr[ , , , 1, ] = return_xyer

# RCP 4.5 Period 
       
  print("Importing RCP 4.5 Case")


  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_45,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    All_Return_Intervals_xyesr_arr[ , , , 2, ] = return_xyer
    

    
# RCP 8.5 Period 
       
  print("Importing RCP 8.5 Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_85,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    All_Return_Intervals_xyesr_arr[ , , , 3, ] = return_xyer

    
    
   remove(LOCA_File_Base,
          LOCA_File_Per1_45,
          LOCA_File_Per1_85,
          LOCA_URL,
          URL_Root_Directory,
          time,
          latitude,
          longitude,
          start_year,
          end_year,
          ensemble_site,
          sumprec_xye,
          meanprec_xy,
          mean_xye,  
          mean_xyde,
          stdv_xye,
          stdvprec_xy,
          return_xyer,
          med_return_xyr,
          sd_return_xyr,
          return_intervals)
   
```


# Pass a Median Filter over Raw Data.

```{r}

  # get thematic mask netcdf file.


ncngp  = nc_open(filename = "//Users/wjc/GitHub/FEW_CROP_MiGRAGTION/arch/NGP_Climate_Zones.nc") 

lon         =  ncvar_get(nc    = ncngp, 
                         varid = "lon")
lat         =  ncvar_get(nc    = ncngp,
                         varid = "lat")

climate_regions  =  array(data     = NA,
                          dim      = c(length(lon),
                                       length(lat)),
                          dimnames = list("lon" = lon,
                                          "lat" = lat))

climate_regions[,]    = ncvar_get(nc    = ncngp,
                               varid = "US_CAN_Zones")

climate_regions = t(apply(X      = climate_regions,
                        MARGIN = 1, 
                        FUN    = rev))


 filled.contour(climate_regions)
 filled.contour(All_Return_Intervals_xyesr_arr[, , 1, 1, 1])
 
 remove(ncngp)
 
Climate_Zones = levels(as.factor(climate_regions))
 
n_zones       = length(Climate_Zones)

 
 
```



# Create Thematic Arrays Zone-Median arrays for Output Array for Reprojected Return Period

Dimensions

  - nx = longtiude 
  - ne = ensembles
  - ns = scenario
  - nr = return invervals 

`
```{r}

# create resampled block


   Cross_Scenario_Reprojected_Returns_zesr_arr = 
            array(data    = NA,
                  dim     = c(length(Climate_Zones),
                              length(Ensemble),
                              length(Scenario3),
                              length(Return_Intervals)),
                  dimnames = list("Climate_Zones"    = Climate_Zones,
                                  "Ensemble"         = Ensemble,
                                  "Scenario"         = Scenario3,
                                  "Return_Intervals" = Return_Intervals))


   Cross_Scenario_Ensemble_Median_Reprojected_Returns_zsr_arr = 
            array(data    = NA,
                  dim     = c(length(Climate_Zones),
                              length(Scenario),
                              length(Return_Intervals)),
                  dimnames = list("Climate_Zones"    = Climate_Zones,
                                  "Scenario"         = Scenario,
                                  "Return_Intervals" = Return_Intervals))
   


```



```{r}


print("Calcuilating Zone and Ensemble Means")

  for (z  in seq(from =                     1,
                 to   = length(Climate_Zones),
                 by   =                     1)) {  # zone loop
    
    
    mask    = climate_regions[,] * NA
    
    mask[climate_regions == as.numeric(Climate_Zones[z])] = 1.0
    
    n_cells = length(which(mask == 1) )
    
    
    print(str_c(" ",
                "Median",
                "  Processing Zone ",
                z,
                "; Zone : ",
                Climate_Zones[z],
                " [",
                (z*100. / length(Climate_Zones) ),
                "%] N=",n_cells,
                sep=""))
    
    for (e in seq(from =                1,
                  to   = length(Ensemble),
                  by   =                1)) {  # ensemble  loop
      
      for (s in seq(from =    1,
                    to   = length(Scenario3),
                    by   = 1)) {  # scenario  loop
        
        for (r in seq(from =    1,
                      to   = length(Return_Intervals),
                      by   = 1)) { #return year loop
          
          subset = All_Return_Intervals_xyesr_arr[,,e,s,r] * mask
          

          Cross_Scenario_Reprojected_Returns_zesr_arr[z,e,s,r] = median(x = subset,
                                                                            na.rm=TRUE)
          }  #return year loop 
        
        } # scenario  loop
      
      } # ensemble  loop
    
    }  # zone loop




```



## Calcualte Inversion Function for Historical

```{r}

print("Resampling Return Years")

n_ens  = length(Ensemble)
n_zone = length(Latitude)

for (k in seq(from =     1,
              to   = n_ens,
              by   =     1))  {  # ensemble loop
  
  print(str_c("Ensemble ",
              Ensemble[k],
              sep = ""))
  
  ensemble_extracted = All_Return_Intervals_xyesr_arr[,,k,, ]

  for (j in seq(from =     1,
                to   = n_zone,
                by   =     1)) {  # Zone loop
    

      print(str_c(" ",
                  Ensemble[k],
                  "  Processing Row ",
                  j,
                  "; Latitude : ",
                  Latitude[j],
                  "°N; [",
                  (j*100. / n_lat ),
                  "%]",
                  sep=""))
      

    
    
    
    
      
     
    for (i in seq(from =     1,
                  to   = n_lon,
                  by   =     1)) {   # longitude loop
      

        if ( !anyNA( c( All_Return_Intervals_xyesr_arr[i,j,k,1, ],
                        All_Return_Intervals_xyesr_arr[i,j,k,2, ],
                        All_Return_Intervals_xyesr_arr[i,j,k,3, ]) ) )  { # NA Loop
            
            # Calcualte Inversion Function for Historical
          
            funcHS = splinefun(x      = All_Return_Intervals_xyesr_arr[i,j,k,1, ],
                               y      = Return_Intervals, 
                               method = "hyman", 
                               ties   = mean)  
             
            # Calcualte Inversion Function for RCP 4.5
          
            func45 = splinefun(x      = All_Return_Intervals_xyesr_arr[i,j,k,2, ],
                               y      = Return_Intervals, 
                               method = "hyman", 
                               ties   = mean)
          
            # Calcualte Inversion Function for RCP 8.5   
  
            func85 = splinefun(x      = All_Return_Intervals_xyesr_arr[i,j,k,3, ],
                               y      = Return_Intervals, 
                               method = "hyman", 
                               ties   = mean)
            
            # Apply Functions to Historical Case 
            
            Cross_Scenario_Reprojected_Returns_xyesr_arr[i,j,k,1,] =
                 func45(All_Return_Intervals_xyesr_arr[i,j,k,1, ])
              
            Cross_Scenario_Reprojected_Returns_xyesr_arr[i,j,k,2,] = 
                 func85(All_Return_Intervals_xyesr_arr[i,j,k,1, ])
   
        } # NA Loop
        
     }  # longitude loop
  
  }  # Zone loop

}  # ensemble loop



#Cross_Scenario_Reprojected_Returns_xyesr_arr = as.numeric(Cross_Scenario_Reprojected_Returns_xyesr_arr)

```



```{r}


  save(period_span,
       base_start,
       base_end,
       per1_start,
       per1_end,
       Latitude,
       Longitude,
       Ensemble,
       Scenario,
       Scenario3,
       All_Return_Intervals_xyesr_arr,
       Cross_Scenario_Reprojected_Returns_xyesr_arr,
       Cross_Scenario_Ensemble_Median_Reprojected_Returns_xysr_arr,
       file =  paste("./LOCA_NGP_Ensemble_Return_Periods_",
                     base_string,
                     "_vs_",
                     per1_string,
                     ".RData",
                     sep="")
       )





```


```{r}



  Resampled_Return_Periods = melt(data = Cross_Scenario_Reprojected_Returns_xyesr_arr,
                                  value.name = "Resampled_Future_Return_Periods")
  remove(Cross_Scenario_Reprojected_Returns_xyesr_arr)


  Resampled_Return_Periods$Ensemble = as.factor(Resampled_Return_Periods$Ensemble)
  Resampled_Return_Periods$Scenario = as.factor(Resampled_Return_Periods$Scenario)

  
  
    save(
       base_start,
       base_end,
       per1_start,
       per1_end,
       period_span,
       Longitude,
       Latitude,
       Ensemble,
       Scenario,
       Return_Intervals,
       Resampled_Return_Periods,
       file =  paste("./LOCA_NGP_Resampled_Return_Periods__",
                     base_start,"-",base_end,
                     "_vs_",
                     per1_start,"-",per1_end,
                     ".RData",
                     sep="")
       )

    remove(Resampled_Return_Periods)

    
    
    
  Median_Resampled_Return_Periods = melt(data = Cross_Scenario_Ensemble_Median_Reprojected_Returns_xysr_arr,
                                         value.name = "Resampled_Future_Return_Periods")  
    
    
    remove(Cross_Scenario_Ensemble_Median_Reprojected_Returns_xysr_arr)
  
      save(
       base_start,
       base_end,
       per1_start,
       per1_end,
       period_span,
       Longitude,
       Latitude,
       Ensemble,
       Scenario,
       Return_Intervals,
       Median_Resampled_Return_Periods,
       file =  paste("./LOCA_NGP_Ensemble-Median_Resampled_Return_Periods_smoothed__",
                     base_start,"-",base_end,
                     "_vs_",
                     per1_start,"-",per1_end,
                     ".RData",
                     sep=""))
      
          remove(Median_Resampled_Return_Periods)

  
```

##################
```{r}


    
         
  All_Return_Intervals = melt(data = All_Return_Intervals_xyesr_arr,
                                  value.name = "Return_Period_Values")

  remove(All_Return_Intervals_xyesr_arr)
  
        

  All_Return_Intervals$Ensemble = as.factor(All_Return_Intervals$Ensemble)
  All_Return_Intervals$Scenario = as.factor(All_Return_Intervals$Scenario)

  
    save(
       base_start,
       base_end,
       per1_start,
       per1_end,
       period_span,
       Longitude,
       Latitude,
       Ensemble,
       Scenario3,
       Return_Intervals,
       All_Return_Intervals,
       file =  paste("./LOCA_NGP_Ensemble_All_121Med_Return_Periods__",
                     base_start,"-",base_end,
                     "_vs_",
                     per1_start,"-",per1_end,
                     ".RData",
                     sep="")
       )


```
=