---
title: "Extreme Precip Analysis or a Grid over a Raster Grid Original Product"
output: html_notebook
---



Processing for Max Precip Daily Extremes Using LOCA Model Gridded Output.

Temperature is tricky since we will be using a Generalized Pareto to target the extreme values and that also requires a threshold to which the method is sensitive.  Temperature complicates this since we are looking at a goodly amount of non-stationarity as we move into the future.

Warning Typos are Legion.



# Load Required Libraries

```{r}
  library(package = "tidyverse")
  library(package = "tidypredict")

  library(package = "lubridate") # processing dates and time
  library(package = "stringr")


  library(package = "reshape2")  # manipulating data frames
  library(package = "extRemes")  # extreme data analysis
  library(package = "abind")
  
  
```

# Available Ensembles

**Table for Available LOCA CMIP5 Point Stations**

CMIP5 Models        | CMIP5 Models          | CMIP5 Models
--------------------|-----------------------|---------------------
ACCESS1.0_r1i1p1    | ACCESS1.3_r1i1p1      | CCSM4_r6i1p1         
CESM1.BGC_r1i1p1    | CESM1.CAM5_r1i1p1     | CMCC.CMS_r1i1p1      
CMCC.CM_r1i1p1      | CNRM.CM5_r1i1p1       | CSIRO.Mk3.6.0_r1i1p1 
CanESM2_r1i1p1      | FGOALS.g2_r1i1p1      | GFDL.CM3_r1i1p1      
GFDL.ESM2G_r1i1p1   | GFDL.ESM2M_r1i1p1     | HadGEM2.AO_r1i1p1    
HadGEM2.CC_r1i1p1   | HadGEM2.ES_r1i1p1     | IPSL.CM5A.LR_r1i1p1  
IPSL.CM5A.MR_r1i1p1 | MIROC.ESM.CHEM_r1i1p1 | MIROC.ESM_r1i1p1     
MIROC5_r1i1p1       | MPI.ESM.LR_r1i1p1     | MPI.ESM.MR_r1i1p1    
MRI.CGCM3_r1i1p1    | NorESM1.M_r1i1p1      | bcc.csm1.1.m_r1i1p1


And we should include our time periods.  Here, we have a baseline period and a future period

```{r}
  period_span     = 30.
  base_start      = 1976              # start year
  base_end        = base_start + period_span-1 # end year  (just before our simulations diverge)
  
  base_string     = str_c(base_start,
                          base_end,
                          sep="-")
  
  
  
  per1_start      = 2035             # start
  per1_end        = per1_start + period_span-1 # end yvar2dear
  
  per1_string     = str_c(per1_start,
                          per1_end,
                          sep="-")
  
  
  Scenario3 = c( str_c("Historical (",
                        base_string,
                        ")",
                        sep = ""),
                
                 str_c("RCP 4.5 (",
                        per1_string,
                        ")",
                        sep = ""),
                
                 str_c("RCP 8.5 (",
                        per1_string,
                       ")",
                        sep = "")
                
                )

  Scenario  = c( str_c("RCP 4.5 (",
                        per1_string,
                        ") - (",
                        base_string,
                        ")",
                        sep = ""),
                
                  str_c("RCP 8.5 (",
                        per1_string,
                       ") - (",
                        base_string,
                        ")",
                        sep = "")
                )
      
```

# Thresholds for Extreme Events

And while I am up here I am defining a "wetting rain" even threshold as any rain event exceeding 0.1" per day

```{r}

  threshold_ptile = 0.85
  
  wetting_rain_treshold = 0.1 * 25.4 # in -> mm (0.1 in = 2.54 )

```

# Data Loading and Pre-Processing

## File Control

```{r}

  URL_Root_Directory = "http://kyrill.ias.sdsmt.edu:8080/thredds/dodsC/LOCA_NGP/climatology/"
  URL_Root_Directory = "/maelstrom2/LOCA_GRIDDED_ENSEMBLES/LOCA_NGP/climatology/"
  
  # get URL files  "return_statistics_NGP_LOCA_rcp45_pr_2035-2064.RData""
  LOCA_File_Base  = str_c(base_string,
                          "/",
                          "DERIVED",
                          "/",
                          "return_statistics_NGP_LOCA_historical_pr_",
                          base_string,
                          ".RData",
                          sep="")

  LOCA_File_Per1_45  = str_c(per1_string,
                             "/",
                             "DERIVED",
                             "/",
                             "return_statistics_NGP_LOCA_rcp45_pr_",
                             per1_string,
                             ".RData",
                             sep="")

  LOCA_File_Per1_85  = str_c(per1_string,
                             "/",
                             "DERIVED",
                             "/",
                             "return_statistics_NGP_LOCA_rcp85_pr_",
                             per1_string,
                             ".RData",
                             sep="")
  

```


## Data Ingest


```{r}

# Base Period 

  print("Importing Historical Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Base,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    base_med_return_xyr_arr = med_return_xyr
    

# RCP 4.5 Period 
       
  print("Importing RCP 4.5 Case")


  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_45,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    per1_85_med_return_xyr_arr  = med_return_xyr
    

    
# RCP 8.5 Period 
       
  print("Importing RCP 8.5 Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_85,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    per1_45_med_return_xyr_arr  = med_return_xyr


# Fix the Dimensions 

  dimnames(base_med_return_xyr_arr)    = list("longitude"        = longitude,
                                              "latitude"         = latitude,
                                              "Return_Intervals" = return_intervals) 
       
  dimnames(per1_45_med_return_xyr_arr) = list("longitude"        = longitude,
                                              "latitude"         = latitude,
                                              "Return_Intervals" = return_intervals)
  
  dimnames(per1_85_med_return_xyr_arr) = list("longitude"        = longitude,
                                              "latitude"         = latitude,
                                              "Return_Intervals" = return_intervals) 
 


  
  Ensemble = ensemble_site
  
  Return_Intervals = return_intervals
    
   
  remove(LOCA_File_Base,
         LOCA_File_Per1_45,
         LOCA_File_Per1_85,
         LOCA_URL,
         URL_Root_Directory)
  
     
   remove(time,
          start_year,
          end_year,
          ensemble_site,
          sumprec_xye,
          meanprec_xy,
          mean_xye,  
          mean_xyde,
          stdv_xye,
          stdvprec_xy,
          return_xyer,
          med_return_xyr,
          sd_return_xyr,
          return_intervals)
   
  
   
  
```

# Create Raster for Output Array for Reprojected Return Period

Dimensions

  - nx = longtiude 
  - ny = latitude
  - ns = scenario
  - nr = return invervals 

```{r}

   cross_scenario_reprojected_median_return_xysr_arr = 
            array(data     = NA,
                  dim      = c(length(longitude),
                               length(latitude),
                               length(Scenario),
                               length(Return_Intervals)),
                  dimnames = list("longitude"                = longitude,
                                  "latitude"                 = latitude,
                                  "Scenario"                 = Scenario,
                                  "Base_Period_Return_Periods" = Return_Intervals)
                  )

```




## Calcualte Inversion Function for Historical

```{r}


n_lat = length(latitude)
n_lon = length(longitude)

for (jj in seq(from =     1,
               to   = n_lat,
               by   =     1)) {  # latitude loop
  
  if ((jj %% 10) == 0) { # in-loop entertainment
  
  print(str_c("Processing Row ",
              jj,
              "; Latitude : ",
              latitude[jj],
              "°N; [",
              (jj*100. / n_lat ),"%]",
              sep=""))
  } # in-loop entertainment
    
   
  for (ii in seq(from =     1,
                 to   = n_lon,
                 by   =     1)) {   # longitude loop
    

      historical_return_values = base_med_return_xyr_arr[ii,jj, ]
      rcp45_return_values      = per1_45_med_return_xyr_arr[ii,jj, ]
      rcp85_return_values      = per1_85_med_return_xyr_arr[ii,jj, ]
      
      if ( !anyNA( c( historical_return_values,
                      rcp45_return_values,
                      rcp85_return_values) ) )  { # NA Loop
          
          # Calcualte Inversion Function for Historical
        
          funcHS = splinefun(x      = historical_return_values,
                             y      = Return_Intervals, 
                             method = "hyman", 
                             ties   = mean)  
           
          # Calcualte Inversion Function for RCP 4.5
        
          func45 = splinefun(x      = rcp45_return_values,
                             y      = Return_Intervals, 
                             method = "hyman", 
                             ties   = mean)
        
          # Calcualte Inversion Function for RCP 8.5   

          func85 = splinefun(x      = rcp85_return_values,
                             y      = Return_Intervals, 
                             method = "hyman", 
                             ties   = mean)
          
          # Apply Functions to Historical Case 
          
          cross_scenario_reprojected_median_return_xysr_arr[ii,jj,1,] =
               func45(historical_return_values)
            
          cross_scenario_reprojected_median_return_xysr_arr[ii,jj,2,] = 
               func85(historical_return_values)
 
      
      } # NA Loop

      
   }  # latitude loop

  
}  # longitude loop


```


```{r}
  # Patch the Resampled Return Periods into a Frame

  Resampled_Return_Periods = melt(data = cross_scenario_reprojected_median_return_xysr_arr,
                                  value.name = "Resampled_Future_Return_Periods")


  # Patch the Median Return Periods into a Frame

  Median_Return_Periods = melt(data       = base_med_return_xyr_arr,
                          varnames   = names(dimnames(base_med_return_xyr_arr)),
                          value.name = "Historical")
   
   per1_med_return      = melt(data      = per1_45_med_return_xyr_arr,
                          varnames   = names(dimnames(per1_45_med_return_xyr_arr)),
                          value.name = "RCP45") 
   
   Median_Return_Periods$RCP45 = per1_med_return$RCP45
   
   per1_med_return = melt(data      = per1_85_med_return_xyr_arr,
                          varnames   = names(dimnames(per1_85_med_return_xyr_arr)),
                          value.name = "RCP85")  
   
   Median_Return_Periods$RCP85 = per1_med_return$RCP85
   
   Median_Return_Periods = gather(data  = Median_Return_Periods,
                          key   = "Scenario",
                          value = "Median_Return_Period_Values",
                          Historical,
                          RCP45,
                          RCP85)
   
   
# Patch the Scenario Names
   
  Median_Return_Periods = Median_Return_Periods %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "Historical", 
                                                   str_c("Historical (",
                                                         base_string, ")",
                                                         sep=""))) %>% 
                         as.data.frame()
  
  Median_Return_Periods = Median_Return_Periods %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "RCP45", 
                                                   str_c("RCP 4.5 (",
                                                         per1_string, ")",
                                                         sep = ""))) %>% 
                         as.data.frame()
  
  Median_Return_Periods = Median_Return_Periods %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "RCP85", 
                                                   str_c("RCP 8.5 (",
                                                         per1_string, ")",
                                                         sep = ""))) %>% 
                         as.data.frame()
  
  Median_Return_Periods$Scenario    = as.factor(Median_Return_Periods$Scenario)
  Resampled_Return_Periods$Scenario = as.factor(Resampled_Return_Periods$Scenario)
  
  Resampled_Return_Periods$Change_in_Return_Period = 
      Resampled_Return_Periods$Resampled_Future_Return_Periods - 
      Resampled_Return_Periods$Base_Period_Return_Periods
  
  
   

  save(period_span,
       base_start,
       base_end,
       per1_start,
       per1_end,
       Median_Return_Periods,
       Resampled_Return_Periods,
       file =  paste("./LOCA_NGP_Ensemble_Median_Return_Periods_",
                     base_string,
                     "_vs_",
                     per1_string,
                     ".RData",
                     sep="")
       )


  remove(per1_med_return,
         per1_45_med_return_xyr_arr,
         per1_85_med_return_xyr_arr,
         base_med_return_xyr_arr,
         cross_scenario_reprojected_median_return_xysr_arr)
  
```



