---
title: "Extreme Precip Analysis or a Single Point over a Raster Grid Original Product"
output: html_notebook
---



Processing for Max Precip Daily Extremes Using LOCA Model Gridded Output.

Temperature is tricky since we will be using a Generalized Pareto to target the extreme values and that also requires a threshold to which the method is sensitive.  Temperature complicates this since we are looking at a goodly amount of non-stationarity as we move into the future.

Warning Typos are Legion.



# Load Required Libraries

```{r}
  library(package = "tidyverse")
  library(package = "tidypredict")

  library(package = "lubridate") # processing dates and time
  library(package = "stringr")


  library(package = "reshape2")  # manipulating data frames
  library(package = "extRemes")  # extreme data analysis
  library(package = "abind")
  
  
```

# Available Ensembles

**Table for Available LOCA CMIP5 Point Stations**

CMIP5 Models        | CMIP5 Models          | CMIP5 Models
--------------------|-----------------------|---------------------
ACCESS1.0_r1i1p1    | ACCESS1.3_r1i1p1      | CCSM4_r6i1p1         
CESM1.BGC_r1i1p1    | CESM1.CAM5_r1i1p1     | CMCC.CMS_r1i1p1      
CMCC.CM_r1i1p1      | CNRM.CM5_r1i1p1       | CSIRO.Mk3.6.0_r1i1p1 
CanESM2_r1i1p1      | FGOALS.g2_r1i1p1      | GFDL.CM3_r1i1p1      
GFDL.ESM2G_r1i1p1   | GFDL.ESM2M_r1i1p1     | HadGEM2.AO_r1i1p1    
HadGEM2.CC_r1i1p1   | HadGEM2.ES_r1i1p1     | IPSL.CM5A.LR_r1i1p1  
IPSL.CM5A.MR_r1i1p1 | MIROC.ESM.CHEM_r1i1p1 | MIROC.ESM_r1i1p1     
MIROC5_r1i1p1       | MPI.ESM.LR_r1i1p1     | MPI.ESM.MR_r1i1p1    
MRI.CGCM3_r1i1p1    | NorESM1.M_r1i1p1      | bcc.csm1.1.m_r1i1p1


And we should include our time periods.  Here, we have a baseline period and a future period

```{r}
  period_span     = 30.
  base_start      = 1976              # start year
  base_end        = base_start + period_span-1 # end year  (just before our simulations diverge)
  
  base_string     = str_c(base_start,
                          base_end,
                          sep="-")
  
  
  
  per1_start      = 2035             # start
  per1_end        = per1_start + period_span-1 # end yvar2dear
  
  per1_string     = str_c(per1_start,
                          per1_end,
                          sep="-")
```

# Thresholds for Extreme Events

And while I am up here I am defining a "wetting rain" even threshold as any rain event exceeding 0.1" per day

```{r}

  threshold_ptile = 0.85
  
  wetting_rain_treshold = 0.1 * 25.4 # in -> mm (0.1 in = 2.54 )

```

# Data Loading and Pre-Processing

## File Control

```{r}

  URL_Root_Directory = "http://kyrill.ias.sdsmt.edu:8080/thredds/dodsC/LOCA_NGP/climatology/"
  URL_Root_Directory = "/maelstrom2/LOCA_GRIDDED_ENSEMBLES/LOCA_NGP/climatology/"
  
  # get URL files  "return_statistics_NGP_LOCA_rcp45_pr_2035-2064.RData""
  LOCA_File_Base  = str_c(base_string,
                          "/",
                          "DERIVED",
                          "/",
                          "return_statistics_NGP_LOCA_historical_pr_",
                          base_string,
                          ".RData",
                          sep="")

  LOCA_File_Per1_45  = str_c(per1_string,
                             "/",
                             "DERIVED",
                             "/",
                             "return_statistics_NGP_LOCA_rcp45_pr_",
                             per1_string,
                             ".RData",
                             sep="")

  LOCA_File_Per1_85  = str_c(per1_string,
                             "/",
                             "DERIVED",
                             "/",
                             "return_statistics_NGP_LOCA_rcp85_pr_",
                             per1_string,
                             ".RData",
                             sep="")
  

```


## Data Ingest


```{r}

# Base Period 

  print("Importing Historical Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Base,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    base_med_return_xyr_arr = med_return_xyr
    

# RCP 4.5 Period 
       
  print("Importing RCP 4.5 Case")


  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_45,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    per1_85_med_return_xyr_arr  = med_return_xyr
    

    
# RCP 8.5 Period 
       
  print("Importing RCP 8.5 Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_85,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    per1_45_med_return_xyr_arr  = med_return_xyr


# Fix the Dimensions 

  dimnames(base_med_return_xyr_arr)    = list("longitude"        = longitude,
                                              "latitude"         = latitude,
                                              "Return_Intervals" = return_intervals) 
       
  dimnames(per1_45_med_return_xyr_arr) = list("longitude"        = longitude,
                                              "latitude"         = latitude,
                                              "Return_Intervals" = return_intervals)
  
  dimnames(per1_85_med_return_xyr_arr) = list("longitude"        = longitude,
                                              "latitude"         = latitude,
                                              "Return_Intervals" = return_intervals) 
 


  
  Ensemble = ensemble_site
    
   
  remove(LOCA_File_Base,
         LOCA_File_Per1_45,
         LOCA_File_Per1_85,
         LOCA_URL,
         URL_Root_Directory)
  
     
   remove(time,
          start_year,
          end_year,
          ensemble_site,
          sumprec_xye,
          meanprec_xy,
          mean_xye,  
          mean_xyde,
          stdv_xye,
          stdvprec_xy,
          return_xyer,
          med_return_xyr,
          sd_return_xyr)
   
  
```

## Melt 3d Arrays to Data Frames (Return Period)

```{r}

# Melt 3d Arrays to Data Frames
   
   median_return   = melt(data       = base_med_return_xyr_arr,
                          varnames   = names(dimnames(base_med_return_xyr_arr)),
                          value.name = "Historical")
   
   per1_med_return = melt(data      = per1_45_med_return_xyr_arr,
                          varnames   = names(dimnames(per1_45_med_return_xyr_arr)),
                          value.name = "RCP45") 
   
   median_return$RCP45 = per1_med_return$RCP45
   
   per1_med_return = melt(data      = per1_85_med_return_xyr_arr,
                          varnames   = names(dimnames(per1_85_med_return_xyr_arr)),
                          value.name = "RCP85")  
   
   median_return$RCP85 = per1_med_return$RCP85
   
   remove(per1_med_return,
          base_med_return_xyr_arr,
          per1_45_med_return_xyr_arr,
          per1_85_med_return_xyr_arr)

```




## Gather Into a Single Data Frame (Median of Ensembles)

```{r}
   
# Gather into a single data frame
   
   median_return = gather(data  = median_return,
                          key   = "Scenario",
                          value = "Median_Return_Period_Values",
                          Historical,
                          RCP45,
                          RCP85)
   
# Patch the Scenario Names
   
  median_return = median_return %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "Historical", 
                                                   str_c("Historical (",
                                                         base_string, ")",
                                                         sep=""))) %>% 
                         as.data.frame()
  
  median_return = median_return %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "RCP45", 
                                                   str_c("RCP 4.5 (",
                                                         per1_string, ")",
                                                         sep = ""))) %>% 
                         as.data.frame()
  
  median_return = median_return %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "RCP85", 
                                                   str_c("RCP 8.5 (",
                                                         per1_string, ")",
                                                         sep = ""))) %>% 
                         as.data.frame()
  
  median_return$Scenario = as.factor(median_return$Scenario)

  print("Listing Available Median Scenarios")
  
  levels(median_return$Scenario)
   
  print("Median Data is Read In and We are Ready to Play")
```





## Calcualte Inversion Function for Historical

```{r}

first_run = TRUE

user_requested_lon = -87.4774
user_requested_lat =  36.6634




target_lon = longitude[which((longitude - user_requested_lon)^2 == min((longitude - user_requested_lon)^2))]
target_lat =  latitude[which((latitude  - user_requested_lat)^2 == min((latitude  - user_requested_lat)^2))]

print(str_c(" Requested Latitude = ",
            user_requested_lat,
            sep = ""))
print(str_c("  Targeted Latitude = ",
            target_lat,
            sep = ""))
print("")

print(str_c("Requested Longitude = ",
            user_requested_lon,
            sep = ""))
print(str_c(" Targeted Longitude = ",
            target_lon,
            sep = ""))

n_i = length(longitude)
n_j = length(latitude)

#for (target_lat in latitude) {  # latitude loop
  
  jj = which(latitude  == target_lat)
 
  print(str_c("Processing Row ",
              target_lat,
              "°N [",
              (jj*100. / n_j ),"%]",
              sep=""))
    
   
#  for (target_lon in longitude) {   # longitude loop
    
      ii = which(latitude  == target_lat)
      
      if ( !anyNA( median_return %>% 
                      filter((longitude == target_lon) &
                             (latitude  == target_lat) ) ))  { # NA Loop
          
         
          # Calcualte Inversion Function for Historical
        
          subset = median_return %>% 
                        filter((longitude == target_lon) &
                               (latitude  == target_lat) &
                               str_detect(string  =     Scenario,
                                          pattern = "Historical")  )
        
          return_period_median_resampled_A = data.frame(Return_Period_Values =
                                                        subset$Median_Return_Period_Values)
          

          funcHS = splinefun(x      = subset$Median_Return_Period_Values,
                             y      = subset$Return_Intervals, 
                             method = "fmm", 
                             ties   = mean)  
           
          # Calcualte Inversion Function for RCP 4.5
        
          subset = median_return %>%
                        filter((longitude == target_lon) &
                               (latitude  == target_lat) &
                               str_detect(string  =     Scenario,
                                          pattern = "RCP 4.5")  )                        
        
          func45 = splinefun(x      = subset$Median_Return_Period_Values,
                             y      = subset$Return_Intervals, 
                             method = "fmm", 
                             ties   = mean)
        
          # Calcualte Inversion Function for RCP 8.5   
           
          subset = median_return %>%
                        filter((longitude == target_lon) &
                               (latitude  == target_lat) &
                               str_detect(string  =     Scenario,
                                          pattern = "RCP 8.5")  )                              
                                       
          func85 = splinefun(x      = subset$Median_Return_Period_Values,
                             y      = subset$Return_Intervals, 
                             method = "fmm", 
                             ties   = mean)
          
          # Apply Functions to Historical Case 
          
        
          
          return_period_median_resampled_A$Historical = 
               funcHS(return_period_median_resampled_A$Return_Period_Values)
          
          return_period_median_resampled_A$RCP45      = 
               func45(return_period_median_resampled_A$Return_Period_Values)
          
          return_period_median_resampled_A$RCP85      = 
               func85(return_period_median_resampled_A$Return_Period_Values)
          
          
          
          
          
          return_period_median_resampled_A   = gather(data  = return_period_median_resampled_A,
                                                     key   = "Scenario",
                                                     value = "Resampled_Return_Periods",
                                                     c("Historical","RCP45","RCP85"))
          
        
          
          return_period_median_resampled_A = return_period_median_resampled_A %>% 
                                 mutate(Scenario = replace(Scenario,
                                                           Scenario == "Historical", 
                                                           str_c("Historical (",
                                                                 base_string, ")",
                                                                 sep=""))) %>% 
                                 as.data.frame()
        
           
          
          return_period_median_resampled_A = return_period_median_resampled_A %>% 
                                 mutate(Scenario = replace(Scenario,
                                                           Scenario == "RCP45", 
                                                           str_c("RCP 4.5 (",
                                                                 per1_string, ")",
                                                                 sep=""))) %>% 
                                 as.data.frame()
          
          
          return_period_median_resampled_A = return_period_median_resampled_A %>% 
                                 mutate(Scenario = replace(Scenario,
                                                           Scenario == "RCP85", 
                                                           str_c("RCP 8.5 (",
                                                                 per1_string, ")",
                                                                 sep=""))) %>% 
                                 as.data.frame()
        
          
          
          if (first_run){  # First Run If-Then-Else Block
            
            return_period_median_resampled = return_period_median_resampled_A
            
            first_run = FALSE
            
          } else {   # First Run If-Then-Else Block
            
            rbind(return_period_median_resampled,
                  return_period_median_resampled_A)
            
          }  # First Run If-Then-Else Block
          
          remove(return_period_median_resampled_A) 
      
      } # NA Loop

      
#   }  # latitude loop
  
  print("-----------------------")
  
#}  # longitude loop




    
```



# Graphics

## Plot Scenario-Resampled Return Periods

```{r}

  ggplot(data = return_period_median_resampled) + 
  
    aes(x     = Return_Period_Values,
        y     = Resampled_Return_Periods,
        color = Scenario) +
  
    theme_bw() +
  
    theme(legend.position       = c(0.02, 0.98),
          legend.justification  = c(0.00, 1.00)) + 
  
    ggtitle(label    =  str_c(target_lat,
                              "°N   ",
                              -target_lon,
                              "°W",
                              sep = ""),
            subtitle = "Changes in Daily Rainfall Return Events") + 
  
    xlab(label = "Daily Rainfall (mm)") +
  
    ylab(label = "Return Periods (years)") +
  
    
  
    scale_colour_manual(values = c("blue", "orange", "red")) + 
  
    scale_x_continuous() + 
  
    ylim( 0, max(return_period_median_resampled$Resampled_Return_Periods)) + 

    geom_line()



```
