---
title: "Resamples Extreme Precip Analysis or a Grid over a Raster Grid Original Product to a Thematically-Aggregated Data Set"
output: html_notebook
---



Processing for Max Precip Daily Extremes Using LOCA Model Gridded Output.

Temperature is tricky since we will be using a Generalized Pareto to target the extreme values and that also requires a threshold to which the method is sensitive.  Temperature complicates this since we are looking at a goodly amount of non-stationarity as we move into the future.

Warning Typos are Legion.



# Load Required Libraries

```{r}
  library(package = "tidyverse")
  library(package = "tidypredict")

  library(package = "ncdf4")


  library(package = "lubridate") # processing dates and time
  library(package = "stringr")


  library(package = "reshape2")  # manipulating data frames
  library(package = "abind")

  library(package = "foreach")
  library(package = "parallel")


  
  
```


```{r}
detectCores(all.tests = FALSE, logical = TRUE)

```

# Available Ensembles

**Table for Available LOCA CMIP5 Point Stations**

CMIP5 Models        | CMIP5 Models          | CMIP5 Models
--------------------|-----------------------|---------------------
ACCESS1.0_r1i1p1    | ACCESS1.3_r1i1p1      | CCSM4_r6i1p1         
CESM1.BGC_r1i1p1    | CESM1.CAM5_r1i1p1     | CMCC.CMS_r1i1p1      
CMCC.CM_r1i1p1      | CNRM.CM5_r1i1p1       | CSIRO.Mk3.6.0_r1i1p1 
CanESM2_r1i1p1      | FGOALS.g2_r1i1p1      | GFDL.CM3_r1i1p1      
GFDL.ESM2G_r1i1p1   | GFDL.ESM2M_r1i1p1     | HadGEM2.AO_r1i1p1    
HadGEM2.CC_r1i1p1   | HadGEM2.ES_r1i1p1     | IPSL.CM5A.LR_r1i1p1  
IPSL.CM5A.MR_r1i1p1 | MIROC.ESM.CHEM_r1i1p1 | MIROC.ESM_r1i1p1     
MIROC5_r1i1p1       | MPI.ESM.LR_r1i1p1     | MPI.ESM.MR_r1i1p1    
MRI.CGCM3_r1i1p1    | NorESM1.M_r1i1p1      | bcc.csm1.1.m_r1i1p1


And we should include our time periods.  Here, we have a baseline period and a future period

```{r}
  period_span     = 30.
  base_start      = 1976              # start year
  base_end        = base_start + period_span-1 # end year  (just before our simulations diverge)
  
  base_string     = str_c(base_start,
                          base_end,
                          sep="-")
  
  
  
  per1_start      = 2035             # start
  per1_end        = per1_start + period_span-1 # end yvar2dear
  
  per1_string     = str_c(per1_start,
                          per1_end,
                          sep="-")
  
  
  Scenario3 = c( str_c("Historical (",
                        base_string,
                        ")",
                        sep = ""),
                
                 str_c("RCP 4.5 (",
                        per1_string,
                        ")",
                        sep = ""),
                
                 str_c("RCP 8.5 (",
                        per1_string,
                       ")",
                        sep = "")
                
                )

  Scenario  = c( str_c("RCP 4.5 (",
                        per1_string,
                        ") - (",
                        base_string,
                        ")",
                        sep = ""),
                
                  str_c("RCP 8.5 (",
                        per1_string,
                       ") - (",
                        base_string,
                        ")",
                        sep = "")
                )
      
```


# Data Loading and Pre-Processing

## File Control

```{r}

  URL_Root_Directory = "http://kyrill.ias.sdsmt.edu:8080/thredds/fileServer/LOCA_NGP/climatology/"
  URL_Root_Directory = "/maelstrom2/LOCA_GRIDDED_ENSEMBLES/LOCA_NGP/climatology/"
  
  # get URL files  "return_statistics_NGP_LOCA_rcp45_pr_2035-2064.RData""
  LOCA_File_Base  = str_c(base_string,
                          "/",
                          "DERIVED",
                          "/",
                          "return_statistics_NGP_LOCA_historical_pr_",
                          base_string,
                          ".RData",
                          sep="")

  LOCA_File_Per1_45  = str_c(per1_string,
                             "/",
                             "DERIVED",
                             "/",
                             "return_statistics_NGP_LOCA_rcp45_pr_",
                             per1_string,
                             ".RData",
                             sep="")

  LOCA_File_Per1_85  = str_c(per1_string,
                             "/",
                             "DERIVED",
                             "/",
                             "return_statistics_NGP_LOCA_rcp85_pr_",
                             per1_string,
                             ".RData",
                             sep="")
  

```


## Data Ingest


```{r}

# Base Period 

  print("Importing Historical Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Base,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
    
    
#   Pull Coordinates
    
    Return_Intervals = return_intervals
    Ensemble         = ensemble_site
    Latitude         = latitude
    Longitude        = longitude
    
    All_Return_Intervals_xyesr_arr = array(data    = NA,
                                           dim     = c(length(Longitude),
                                                       length(Latitude),
                                                       length(Ensemble),
                                                       length(Scenario3),
                                                       length(Return_Intervals)),
                                           dimnames = list("Longitude"        = Longitude,
                                                           "Latitude"         = Latitude,
                                                           "Ensemble"         = Ensemble,
                                                           "Scenario"         = Scenario3,
                                                           "Return_Intervals" = Return_Intervals))
    
  
    
    All_Return_Intervals_xyesr_arr[ , , , 1, ] = return_xyer

# RCP 4.5 Period 
       
  print("Importing RCP 4.5 Case")


  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_45,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    All_Return_Intervals_xyesr_arr[ , , , 2, ] = return_xyer
    

    
# RCP 8.5 Period 
       
  print("Importing RCP 8.5 Case")

  LOCA_URL        = str_c(URL_Root_Directory,
                          LOCA_File_Per1_85,
                          sep="")
  
  
   
    load(file = (LOCA_URL)) 
  
    All_Return_Intervals_xyesr_arr[ , , , 3, ] = return_xyer

    
    
   remove(LOCA_File_Base,
          LOCA_File_Per1_45,
          LOCA_File_Per1_85,
          LOCA_URL,
          URL_Root_Directory,
          time,
          latitude,
          longitude,
          start_year,
          end_year,
          ensemble_site,
          sumprec_xye,
          meanprec_xy,
          mean_xye,  
          mean_xyde,
          stdv_xye,
          stdvprec_xy,
          return_xyer,
          med_return_xyr,
          sd_return_xyr,
          return_intervals)
   
```


# Pass a Median Filter over Raw Data.

```{r}

  # get thematic mask netcdf file.


ncngp  = nc_open(filename = "//Users/wjc/GitHub/FEW_CROP_MiGRAGTION/arch/NGP_Climate_Zones.nc") 

lon         =  ncvar_get(nc    = ncngp, 
                         varid = "lon")
lat         =  ncvar_get(nc    = ncngp,
                         varid = "lat")

climate_regions  =  array(data     = NA,
                          dim      = c(length(lon),
                                       length(lat)),
                          dimnames = list("lon" = lon,
                                          "lat" = lat))

climate_regions[,]    = ncvar_get(nc    = ncngp,
                               varid = "US_CAN_Zones")

climate_regions = t(apply(X      = climate_regions,
                        MARGIN = 1, 
                        FUN    = rev))


 filled.contour(climate_regions)
 filled.contour(All_Return_Intervals_xyesr_arr[, , 1, 1, 1])
 
 remove(ncngp)
 
Climate_Zones = levels(as.factor(climate_regions))
 
n_zones       = length(Climate_Zones)

 
 
```



# Create Thematic Arrays Zone-Median arrays for Output Array for Reprojected Return Period

Dimensions

  - nx = longtiude 
  - ne = ensembles
  - ns = scenario
  - nr = return invervals 

`
```{r}

# create resampled block

  Statistic = c("P000",
                "P025",
                "P050",
                "P075",
                "P100",
                "Mean")


   Cross_Scenario_Reprojected_Returns_Stats_zesr_arr = 
            array(data    = NA,
                  dim     = c(length(Climate_Zones),
                              length(Ensemble),
                              length(Scenario3),
                              length(Return_Intervals),
                              length(Statistic)),
                  dimnames = list("Climate_Zones"    = Climate_Zones,
                                  "Ensemble"         = Ensemble,
                                  "Scenario"         = Scenario3,
                                  "Return_Intervals" = Return_Intervals,
                                  "Statistic"        = Statistic)
                  )



    
       
```



```{r}


print("Calcuilating Zone and Ensemble Means")

  for (z  in seq(from =                     1,
                 to   = length(Climate_Zones),
                 by   =                     1)) {  # zone loop
    
    
    mask    = climate_regions[,] * NA
    
    mask[climate_regions == as.numeric(Climate_Zones[z])] = 1.0
    
    n_cells = length(which(mask == 1) )
    
    
    print(str_c(" ",
                "Median",
                "  Processing Zone ",
                z,
                "; Zone : ",
                Climate_Zones[z],
                " [",
                (z*100. / length(Climate_Zones) ),
                "%] N=",n_cells,
                sep=""))
    
    for (e in seq(from =                1,
                  to   = length(Ensemble),
                  by   =                1)) {  # ensemble  loop
      
      for (s in seq(from =    1,
                    to   = length(Scenario3),
                    by   = 1)) {  # scenario  loop
        
        for (r in seq(from =    1,
                      to   = length(Return_Intervals),
                      by   = 1)) { #return year loop
          
          subset = All_Return_Intervals_xyesr_arr[,,e,s,r] * mask


          
          
          Cross_Scenario_Reprojected_Returns_Stats_zesr_arr[z,e,s,r,1] = min(x = subset,
                                                                               na.rm=TRUE)
          
          Cross_Scenario_Reprojected_Returns_Stats_zesr_arr[z,e,s,r,2] = quantile(x = subset, probs=0.25,
                                                                          na.rm=TRUE)

          Cross_Scenario_Reprojected_Returns_Stats_zesr_arr[z,e,s,r,3] = median(x = subset,
                                                                             na.rm=TRUE)
          
          
          Cross_Scenario_Reprojected_Returns_Stats_zesr_arr[z,e,s,r,4] = quantile(x = subset, probs=0.75,
                                                                          na.rm=TRUE)
                    

          
          
          Cross_Scenario_Reprojected_Returns_Stats_zesr_arr[z,e,s,r,5] = max(x = subset,
                                                                          na.rm=TRUE)          
                    
          Cross_Scenario_Reprojected_Returns_Stats_zesr_arr[z,e,s,r,6] = mean(x = subset,
                                                                               na.rm=TRUE)
          
          }  #return year loop 
        
        } # scenario  loop
      
      } # ensemble  loop
    
    }  # zone loop



```






```{r}


  Ensemble_Return_Periods_By_Climate_Zone_Stats = melt(data = Cross_Scenario_Reprojected_Returns_Stats_zesr_arr,
                                                        value.name = "Zone_Median_Return_Periods")

  Ensemble_Return_Periods_By_Climate_Zone_Stats$Return_Intervals = as.factor(Ensemble_Return_Periods_By_Climate_Zone_Stats$Return_Intervals)
  
  
    Ensemble_Return_Periods_By_Climate_Zone_Stats$Statistic = as.factor(Ensemble_Return_Periods_By_Climate_Zone_Stats$Statistic) 

  save(period_span,
       base_start,
       base_end,
       per1_start,
       per1_end,
       Climate_Zones,
       Return_Intervals,
       Ensemble,
       Scenario3,
       Ensemble_Return_Periods_By_Climate_Zone_Stats,
       file =  paste("./LOCA_NGP_Ensemble_Return_Periods_by_Climate_Zone_Statistics_",
                     base_string,
                     "_vs_",
                     per1_string,
                     ".RData",
                     sep="")
       )





```


=