---
title: "Extreme Precip Analysis or a Grid over a Raster Grid Original Product"
output: html_notebook
---



Processing for Max Precip Daily Extremes Using LOCA Model Gridded Output.

Temperature is tricky since we will be using a Generalized Pareto to target the extreme values and that also requires a threshold to which the method is sensitive.  Temperature complicates this since we are looking at a goodly amount of non-stationarity as we move into the future.

Warning Typos are Legion.



# Load Required Libraries

```{r}
  library(package = "tidyverse")
  library(package = "tidypredict")

  library(package = "ncdf4")


  library(package = "lubridate") # processing dates and time
  library(package = "stringr")


  library(package = "reshape2")  # manipulating data frames
  library(package = "abind")

  library(package = "foreach")
  library(package = "parallel")


  
  
```


```{r}
detectCores(all.tests = FALSE, logical = TRUE)

```

# Available Ensembles

**Table for Available LOCA CMIP5 Point Stations**

CMIP5 Models        | CMIP5 Models          | CMIP5 Models
--------------------|-----------------------|---------------------
ACCESS1.0_r1i1p1    | ACCESS1.3_r1i1p1      | CCSM4_r6i1p1         
CESM1.BGC_r1i1p1    | CESM1.CAM5_r1i1p1     | CMCC.CMS_r1i1p1      
CMCC.CM_r1i1p1      | CNRM.CM5_r1i1p1       | CSIRO.Mk3.6.0_r1i1p1 
CanESM2_r1i1p1      | FGOALS.g2_r1i1p1      | GFDL.CM3_r1i1p1      
GFDL.ESM2G_r1i1p1   | GFDL.ESM2M_r1i1p1     | HadGEM2.AO_r1i1p1    
HadGEM2.CC_r1i1p1   | HadGEM2.ES_r1i1p1     | IPSL.CM5A.LR_r1i1p1  
IPSL.CM5A.MR_r1i1p1 | MIROC.ESM.CHEM_r1i1p1 | MIROC.ESM_r1i1p1     
MIROC5_r1i1p1       | MPI.ESM.LR_r1i1p1     | MPI.ESM.MR_r1i1p1    
MRI.CGCM3_r1i1p1    | NorESM1.M_r1i1p1      | bcc.csm1.1.m_r1i1p1


And we should include our time periods.  Here, we have a baseline period and a future period

# Data Loading and Pre-Processing

## File Control

```{r}


  
  # get URL files  "return_statistics_NGP_LOCA_rcp45_pr_2035-2064.RData""
  Return_Input_File  = "./LOCA_NGP_Ensemble_Return_Periods_merge_all_ens_by_Climate_Zone_Statistics_1976-2005_vs_2035-2064.RData"
  

```


## Data Ingest


```{r}

# Base Period 


  
   
  load(file = Return_Input_File) 

  load(file = "./Climate_Zones_Name_LUT.Rdata")

  base_string     = str_c(base_start,
                          base_end,
                          sep="-")
  
  per1_string     = str_c(per1_start,
                          per1_end,
                          sep="-")
    
  Scenario  = c( str_c("RCP 4.5 (",
                        per1_string,
                        ") - (",
                        base_string,
                        ")",
                        sep = ""),
                
                  str_c("RCP 8.5 (",
                        per1_string,
                       ") - (",
                        base_string,
                        ")",
                        sep = "")
                )
  
  Climate_Zones = unique(Ensemble_Return_Periods_By_Climate_Zone_Stats_All_Ens$Climate_Zones)
  
  
  Ensemble_Return_Periods_By_Climate_Zone_Stats_All_Ens
```


# Create Thematic Arrays Zone-Median arrays for Output Array for Reprojected Return Period

Dimensions

  - nx = longtiude 
  - ne = ensembles
  - ns = scenario
  - nr = return invervals 

`
```{r}

# create resampled block

  z=1

  for (z  in seq(from =                     1,
                 to   = length(Climate_Zones),
                 by   =                     1)) {  # zone loop

    
    
    Target_Zone = Climate_Zones[z]
        
    subset_H = Ensemble_Return_Periods_By_Climate_Zone_Stats_All_Ens %>%
                 filter( (Climate_Zones == Target_Zone) &
                         str_detect(string  = Scenario,
                                    pattern = "Historical") )
    
    subset_8 = Ensemble_Return_Periods_By_Climate_Zone_Stats_All_Ens %>%
                 filter( (Climate_Zones == Target_Zone) &
                         str_detect(string  = Scenario,
                                    pattern = "RCP 8.5") )
    
    subset_4 = Ensemble_Return_Periods_By_Climate_Zone_Stats_All_Ens %>%
                 filter( (Climate_Zones == Target_Zone) &
                         str_detect(string  = Scenario,
                                    pattern = "RCP 4.5") )
    
    if ( nrow(subset_H) > 0 )  { # NA Loop
 
    
      # Calculate Inversion Function for Historical

      funcHS = splinefun(x      = subset_H$Mean,
                         y      = subset_H$Return_Intervals, 
                         method = "hyman", 
                         ties   = mean)  
     
      # Calcualte Inversion Function for RCP 4.5
  
      func45 = splinefun(x      = subset_4$Mean,
                         y      = subset_4$Return_Intervals, 
                         method = "hyman", 
                         ties   = mean)
  
      # Calcualte Inversion Function for RCP 8.5   

      func85 = splinefun(x      = subset_8$Mean,
                         y      = subset_8$Return_Intervals, 
                         method = "hyman", 
                         ties   = mean)
    
      # Apply RCP 4.5 - Hist
    
      temp = data.frame(Climate_Zones      = Target_Zone, 
                        Historical_Return_Intervals   = Return_Intervals,
                        Max_Daily_Rainfall = subset_H$Mean,
                        RCP45              = func45(subset_H$Mean),
                        RCP85              = func85(subset_H$Mean))
    
      if (z == 1) {
      
        
        Resampled_Climate_Zone_Returns = temp
      
      } else {
        
        Resampled_Climate_Zone_Returns =  rbind(Resampled_Climate_Zone_Returns,
                                                temp)
      }
    

      remove(temp)

      remove(func85)
      remove(func45)
      remove(funcHS)
      
    } # NA Loop
    
    print(str_c(" ",
            "  Processing Zone ",
            z,
            "; Zone : ",
            Target_Zone,
            " [",
            (z*100. / length(Climate_Zones) ),
            "%] N=",nrow(subset_H),
            sep=""))
    
    
    remove(subset_4)
    remove(subset_8)
    remove(subset_H)
    
    
    
    
  }  # climate zone loop

  
  Resampled_Climate_Zone_Returns = gather(data = Resampled_Climate_Zone_Returns,
                                          key  = "Scenario",
                                          value = "Future_Return_Intervals",
                                          RCP45,
                                          RCP85)

    Scenario2 = Scenario
  
    Resampled_Climate_Zone_Returns = Resampled_Climate_Zone_Returns %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "RCP45", 
                                                   Scenario2[1])) %>% 
                         as.data.frame()


    Resampled_Climate_Zone_Returns = Resampled_Climate_Zone_Returns %>% 
                         mutate(Scenario = replace(Scenario, 
                                                   Scenario == "RCP85", 
                                                   Scenario2[2])) %>% 
                         as.data.frame()
    
    

```






```{r}



    Resampled_Climate_Zone_Returns = Resampled_Climate_Zone_Returns %>% 
                         mutate(Future_Return_Intervals = replace(Future_Return_Intervals, 
                                                          Future_Return_Intervals <=0, 
                                                          NA)) %>% 
                         as.data.frame()
    

    Resampled_Climate_Zone_Returns$Climate_Zones = as.factor(Resampled_Climate_Zone_Returns$Climate_Zones)
    Resampled_Climate_Zone_Returns$Historical_Return_Intervals = as.factor(Resampled_Climate_Zone_Returns$Historical_Return_Intervals)
    Resampled_Climate_Zone_Returns$Scenario = as.factor(Resampled_Climate_Zone_Returns$Scenario)

  

    save(
       base_start,
       base_end,
       per1_start,
       per1_end,
       period_span,
       Climate_Zones,
       Ensemble,
       Scenario,
       Return_Intervals,
       Resampled_Climate_Zone_Returns,
       file =  paste("./LOCA_NGP_Resampled_Return_Periods_Climate_Zones_",
                     base_start,"-",base_end,
                     "_vs_",
                     per1_start,"-",per1_end,
                     ".RData",
                     sep="")
       )


    
```


```{r}

```




