---
title: "Mapping LOCA for USDA Growing Zones"
output: html_notebook
---

# 1 Introduction

This script processes annual LOCA and GHCN temperature and rainfall.

The climate model output comes from the USGS [LOCA](http://loca.ucsd.edu) Downscaled Analyses from the CMIP5 climate model runs from 27 models.  (There were a total of 32 models used for these analyses but we are only using the model runs that include maximum/miniumum temperature, and precipitation.)

**Table for Available LOCA CMIP5 Point Stations**

CMIP5 Models        | CMIP5 Models          | CMIP5 Models
--------------------|-----------------------|---------------------
ACCESS1.0_r1i1p1    | ACCESS1.3_r1i1p1      | CCSM4_r6i1p1         
CESM1.BGC_r1i1p1    | CESM1.CAM5_r1i1p1     | CMCC.CMS_r1i1p1      
CMCC.CM_r1i1p1      | CNRM.CM5_r1i1p1       | CSIRO.Mk3.6.0_r1i1p1
CanESM2_r1i1p1      | FGOALS.g2_r1i1p1      | GFDL.CM3_r1i1p1      
GFDL.ESM2G_r1i1p1   | GFDL.ESM2M_r1i1p1     | HadGEM2.AO_r1i1p1    
HadGEM2.CC_r1i1p1   | HadGEM2.ES_r1i1p1     | IPSL.CM5A.LR_r1i1p1  
IPSL.CM5A.MR_r1i1p1 | MIROC.ESM.CHEM_r1i1p1 | MIROC.ESM_r1i1p1     
MIROC5_r1i1p1       | MPI.ESM.LR_r1i1p1     | MPI.ESM.MR_r1i1p1    
MRI.CGCM3_r1i1p1    | NorESM1.M_r1i1p1      | bcc.csm1.1.m_r1i1p1

The result is an *ensemble* of 27 independant virtual earths for our simulation period from 1950 to 2099.

The simulations further diverge at 2006 where our ensemble members separate into two scenarios.

* **RCP 4.5**: A simulation targeting a total amount of 4.5 W m^2^ addition greenhouse forcing
* **RCP 8.5**: A simulation targeting a total amount of 8.5 W m^2^ addition greenhouse forcing

Together, we use these two future cases to reprsent upper and lower bounds of likely expected future climates depending on current and projected greenhouse gas emissions. (The "RPC" stands for ["Representative Pathway Concentration"](https://link.springer.com/article/10.1007/s10584-011-0148-z))



# 2  Libraries

* Install Us First
  + [tidyverse](https://www.tidyverse.org) : Set of commonly-used Data Science packages for R that it can install and load all at once. In the long-run you probably also want to install the tidyverse package suite anyway. For this exercise this will include...
    - [gplot2](https://ggplot2.tidyverse.org) : Create Elegant Data Visualizations Using the Grammar of Graphics
    - [tidyr](https://tidyr.tidyverse.org) : tools for shepherding data in data frames.
    - [stringr](https://stringr.tidyverse.org) : Simple, Consistent Wrappers for Common String Operations


  + Additoinal Tidyverse Resources
    - [lubridate](http://lubridate.tidyverse.org) : Date-Time Control


  + NetCDF Resource
    - [ncdf4](https://www.rdocumentation.org/packages/ncdf4/versions/1.16) : NetCDF and OPEnDAP Support
    - [ncdf4.helpers](https://www.rdocumentation.org/packages/ncdf4.helpers/versions/0.3-3) : Additional tools to work with NetCDF & OPEeNDAP files

  + Mapping Resources
    [maps](https://www.rdocumentation.org/packages/maps/versions/3.3.0) : Display of maps. Projection code and larger maps are in separate packages ('mapproj' and 'mapdata').
    [mapproj](https://www.rdocumentation.org/packages/mapproj/versions/1.2.6) : Converts latitude/longitude into projected coordinates.
    [sf](https://www.rdocumentation.org/packages/sf/versions/0.6-3) : Support for simple features, a standardized way to encode spatial vector data

```{r}

# Libraries

  # Tidyverse Packates

  library(package = "tidyverse") # bulk frequently-used tidyverse packages
  library(package = "lubridate") # tidyverse date-time support'


  library(package = "PCICt") # Implementation of POSIXct Work-Alike for 365 and 360 Day Calendars


  # Mapping  support
  library(package = "maps")
  library(package = "mapproj")
  library(package = "rgdal")
  library(package = "rgeos")

  # array -> dataframe
  library(package = "reshape2")



  # netcdf
  library(package = "ncdf4")
  library(package = "ncdf4.helpers")

```





# 3 LOCA Files, Directory and OPeNDAP URLS (includes coordinate extraction)

```{r}

  # file

  root_url = "/maelstrom2/LOCA_GRIDDED_ENSEMBLES/LOCA_NGP/climatology/ANNUAL/tasmin/"
  root_url = "~/GitHub/FEW_CROP_MIGRAGTION/"

  file_url = str_c(root_url,
                   "LOCA_NGP_tasmin_allensembles_allscenarios_2006-2099_30Y_MEAN_ANNUAL_MIN_05Y_PRODUCT_INTERVAL.RData",
                   str_c = "")

  load(file_url)

  longitude = unique(tmin$lon)
  latitude  = unique(tmin$lat)

  tmin$Min_Temp = (tmin$Min_Temp * 9./5.) + 32.

  decade_strings = unique(tmin$time_bnds)

  remove(file_url)
  remove(root_url)
```

# 4 Import Mask Data

```{r}

  # masking information


  load( file = "./NGP_MASKS.Rdata" )

  ngp_regions$Ecoregions_Mask[is.na(ngp_regions$Ecoregions_Mask) |
                              (ngp_regions$Ecoregions_Mask == 1)  ] = 0

  ngp_regions$Ecoregions_Mask[ngp_regions$Ecoregions_Mask >1] = 1

  lat_lon_string_mask = ngp_regions %>% filter(Ecoregions_Mask > 0)
  lat_lon_string_mask = lat_lon_string_mask$lat_lon_string_mask


  lon = ngp_regions$lon_mask

  lat = ngp_regions$lat_mask


```




# 5 Import Mapping Data

US and Lake Data is Available in R.  Canadian Data isn't so we have a custom polygon library

```{r}
load(file = "./canadian_provinces.Rdata")
# inported field is canada_data = canadian_provinces


usa_data   = map_data(map="state",
                      xlim=c(-114.2812, -86.21875),  
                      ylim=c (33.96875,  52.78125) )
water_data = map_data(map="lakes",
                      xlim=c(-114.2812,-86.21875),  
                      ylim=c( 33.96875, 52.78125) )

```

#6 Color Table

```{r}

classes_table = c("1a",
                  "1b",
                  "2a",
                  "2b",
                  "3a",
                  "3b",
                  "4a",
                  "4b",
                  "5a",
                  "5b",
                  "6a",
                  "6b",
                  "7a",
                  "7b",
                  "8a",
                  "8b",
                  "9a",
                  "9b",
                  "10a",
                  "10b",
                  "11a",
                  "11b",
                  "12a",
                  "12b",
                  "13a",
                  "13b" )                



class2L_cols  = c("#D1D0FE", #  "1a",
                  "#BBBCF0", #  "1b",
                  "#A1A1D4", #  "2a",
                  "#E8A5E8", #  "2b",
                  "#DB86E9", #  "3a",
                  "#C772D6", #  "3b",
                  "#9C61FF", #  "4a",
                  "#506AEB", #  "4b",
                  "#6797FF", #  "5a",
                  "#54C2DD", #  "5b",
                  "#3FB13F", #  "6a",
                  "#6DBE4C", #  "6b",
                  "#A2D05F", #  "7a",
                  "#C8D666", #  "7b",
                  "#E9D479", #  "8a",
                  "#E7C44D", #  "8b",
                  "#D6AD45", #  "9a",
                  "#F5AD6D", #  "9b",
                  "#E89130", #  "10a",
                  "#E26D1C", #  "10b",
                  "#E44B1D", #  "11a",
                  "#E57959", #  "11b",
                  "#CE4F46", #  "12a",
                  "#AD4624", #  "12b",
                  "#A16529", #  "13a",
                  "#BA8554") #  "13b"      



class2L_hi = seq(from = -55,
                 to   =  70,
                 by   =   5)

class2L_lo = class2L_hi - 5.


class2L_cols = as.array(class2L_cols)
names(class2L_cols) = as_factor(classes_table)

```


# 7 Loop Through Decades

#7a Minimum Temperatures for Ensembles and Means

```{r}
decade_string = decade_strings[6]
for (decade_string in decade_strings)
{

        tmin_sub  = tmin %>% filter(time_bnds == decade_string)

        tmin_ens_mean = tmin_sub %>% group_by(lat,
                                              lon,
                                              scenario) %>%
                                     summarize(Min_Temp = mean(Min_Temp))

        mymap = ggplot(data   =   tmin_sub)  +

                       aes(x     = lon,
                           y     = lat) +

                       facet_grid(cols = vars(scenario)) +

                       theme_bw() +
                       theme(strip.background = element_rect(fill=NA))+

                       labs(title    = "LOCA Climate Ensemble Analyses",
                            subtitle = str_c(decade_string,
                                             "Mean Annual Extreme Min Temperature",
                                             sep = " ")) +

                       xlab(label = "Longitude") +

                       ylab(label = "Latitude") +

                       coord_map(xlim=c(min(longitude), max(longitude)),
                                 ylim=c(min(latitude), max(latitude)) )  +

                       geom_polygon(data    = usa_data,
                                    mapping = aes(x     =   long,
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA,
                                    color = "black")  +

                      geom_polygon(data  = water_data,
                                    mapping = aes(x     =   long,
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA,
                                    color = "black")  +

                      geom_polygon(data  = canadian_provinces,
                                    mapping = aes(x     =   long,
                                                  y     =    lat,
                                                  color =  Min_Temp,
                                                  group =  group),
                                    fill  = NA,
                                    color = "black")  +

                      geom_contour(data    = ngp_regions,
                                    alpha   =  1,
                                    color   = c("darkgrey"),
                                    size    = 1.1,
                                    breaks  = c( 1),
                                    mapping = aes(x     = lon_mask,
                                                  y     = lat_mask,
                                                  group = NA,
                                                  color = NA,
                                                  z     = Ecoregions_Mask)) +

                       scale_color_gradientn(breaks = class2L_hi,
                                             colors = class2L_cols,
                                             minor_breaks = NULL,
                                             limits       = c(min(class2L_hi)-2.5,
                                                              max(class2L_hi))+2.5,
                                             guide  = guide_colourbar(title     = "°F",
                                                                      barheight = 15)) +
                        geom_contour(data   = tmin_sub,
                                    alpha  = 0.1,
                                    mapping = aes(x     = lon,
                                                  y     = lat,
                                                  color = ..level..,
                                                  group = ensemble,
                                                  z     = Min_Temp)) +
                        geom_contour(data     = tmin_ens_mean,
                                     alpha   = 1,
                                     mapping = aes(x     = lon,
                                                   y     = lat,
                                                   color = ..level..,
                                                   z     = Min_Temp))


        print(mymap)


        ggsave(filename = str_c("./",
                                decade_string,
                                "_LOCA_MINTEMPS_MAP.png",
                                sep = ""),
               plot = mymap,
               device = "png",
             width = 10.5, height = 5.75)


}          


```






#7b USDA Growing Zones for the Ensemble Mean



```{r}

decade_string = decade_strings[9]
for (decade_string in decade_strings)
{

        tmin_sub  = tmin %>% filter(time_bnds == decade_string)

        tmin_ens_mean = tmin_sub %>% group_by(lat,
                                              lon,
                                              scenario) %>%
                                     summarize(Min_Temp = mean(Min_Temp))

        tmin_ens_mean$Zone = "NA"


        for (cat in 1:length(classes_table)) {

          tmin_ens_mean$Zone[(tmin_ens_mean$Min_Temp >  class2L_lo[cat]) &
                             (tmin_ens_mean$Min_Temp <= class2L_hi[cat])]  = classes_table[cat]

        }
        tmin_ens_mean$Zone[tmin_ens_mean$Zone == "NA"] = NA

        tmin_ens_mean$Zone = as_factor(tmin_ens_mean$Zone)


        mymap = ggplot(data   =   tmin_ens_mean)  +

               aes(x     = lon,
                   y     = lat) +

               facet_grid(cols = vars(scenario)) +

               theme_bw() +

               theme(strip.background = element_rect(fill=NA),
                     aspect.ratio     = 1)+

               labs(title    = "LOCA Climate Ensemble Analyses",
                    subtitle = str_c(decade_string,
                                     "Ensemble Mean USDA Plant Heartiness Zones",
                                     sep = " ")) +

               xlab(label = "Longitude") +

               ylab(label = "Latitude") +

               guides(fill=guide_legend(ncol=2)) +



               coord_cartesian(xlim=c(min(longitude), max(longitude)),
                               ylim=c(min(latitude),  max(latitude)) )  +

              # xlim(min(Longitude), max(Longitude)) +
              # ylim(min(Latitude), max(Latitude)) +

               scale_fill_manual(values = class2L_cols,
                                 breaks = classes_table,
                                 drop   = TRUE) +

               geom_raster(data = tmin_ens_mean,
                           mapping = aes(x     = lon,
                                         y     = lat,
                                         fill = Zone)) +

               geom_polygon(data  = usa_data,
                                    mapping = aes(x     =   long,
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA,
                                    color = "black")  +

               geom_polygon(data  = water_data,
                                    mapping = aes(x     =   long,
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA,
                                    color = "black")  +

               geom_polygon(data  = canadian_provinces,
                                    mapping = aes(x     =   long,
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA,
                                    color = "black")



        print(mymap)


        ggsave(filename = str_c("./",
                                decade_string,
                                "_LOCA_GROWING_ZONES_MAP.png",
                                sep = ""),
               plot = mymap,
               device = "png",
             width = 10.5, height = 5.75)


}          


```
