---
title: "Plot Return Comparisons"
output: html_notebook
---


# 1 Introduction

This script processes annual LOCA and GHCN temperature and rainfall.

The climate model output comes from the USGS [LOCA](http://loca.ucsd.edu) Downscaled Analyses from the CMIP5 climate model runs from 27 models.  (There were a total of 32 models used for these analyses but we are only using the model runs that include maximum/miniumum temperature, and precipitation.)

**Table for Available LOCA CMIP5 Point Stations**

CMIP5 Models        | CMIP5 Models          | CMIP5 Models
--------------------|-----------------------|---------------------
ACCESS1.0_r1i1p1    | ACCESS1.3_r1i1p1      | CCSM4_r6i1p1         
CESM1.BGC_r1i1p1    | CESM1.CAM5_r1i1p1     | CMCC.CMS_r1i1p1      
CMCC.CM_r1i1p1      | CNRM.CM5_r1i1p1       | CSIRO.Mk3.6.0_r1i1p1 
CanESM2_r1i1p1      | FGOALS.g2_r1i1p1      | GFDL.CM3_r1i1p1      
GFDL.ESM2G_r1i1p1   | GFDL.ESM2M_r1i1p1     | HadGEM2.AO_r1i1p1    
HadGEM2.CC_r1i1p1   | HadGEM2.ES_r1i1p1     | IPSL.CM5A.LR_r1i1p1  
IPSL.CM5A.MR_r1i1p1 | MIROC.ESM.CHEM_r1i1p1 | MIROC.ESM_r1i1p1     
MIROC5_r1i1p1       | MPI.ESM.LR_r1i1p1     | MPI.ESM.MR_r1i1p1    
MRI.CGCM3_r1i1p1    | NorESM1.M_r1i1p1      | bcc.csm1.1.m_r1i1p1

The result is an *ensemble* of 27 independant virtual earths for our simulation period from 1950 to 2099.

The simulations further diverge at 2006 where our model runs separate into two scenarios.

* **RCP 4.5**: A simulation targeting a total amount of 4.5 W m^2^ addition greenhouse forcing 
* **RCP 8.5**: A simulation targeting a total amount of 8.5 W m^2^ addition greenhouse forcing 

Together, we use these two future cases to reprsent upper and lower bounds of likely expected future climates depending on current and projected greenhouse gas emissions. (The "RPC" stands for ["Representative Pathway Concentration"](https://link.springer.com/article/10.1007/s10584-011-0148-z))



# 2  Libraries

* Install Us First
  + [tidyverse](https://www.tidyverse.org) : Set of commonly-used Data Science packages for R that it can install and load all at once. In the long-run you probably also want to install the tidyverse package suite anyway. For this exercise this will include...
    - [gplot2](https://ggplot2.tidyverse.org) : Create Elegant Data Visualizations Using the Grammar of Graphics
    - [tidyr](https://tidyr.tidyverse.org) : tools for shepherding data in data frames.
    - [stringr](https://stringr.tidyverse.org) : Simple, Consistent Wrappers for Common String Operations


  + Additoinal Tidyverse Resources
    - [lubridate](http://lubridate.tidyverse.org) : Date-Time Control


  + NetCDF Resource
    - [ncdf4](https://www.rdocumentation.org/packages/ncdf4/versions/1.16) : NetCDF and OPEnDAP Support
    - [ncdf4.helpers](https://www.rdocumentation.org/packages/ncdf4.helpers/versions/0.3-3) : Additional tools to work with NetCDF & OPEeNDAP files
    
  + Mapping Resources
    [maps](https://www.rdocumentation.org/packages/maps/versions/3.3.0) : Display of maps. Projection code and larger maps are in separate packages ('mapproj' and 'mapdata').
    [mapproj](https://www.rdocumentation.org/packages/mapproj/versions/1.2.6) : Converts latitude/longitude into projected coordinates.
    [sf](https://www.rdocumentation.org/packages/sf/versions/0.6-3) : Support for simple features, a standardized way to encode spatial vector data
    
    [scales]() : Color Scale Support
  
```{r}

# Libraries

  # Tidyverse Packates

  library(package = "tidyverse") # bulk frequently-used tidyverse packages
  library(package = "lubridate") # tidyverse date-time support'

  library(package = "RColorBrewer")


  library(package = "PCICt") # Implementation of POSIXct Work-Alike for 365 and 360 Day Calendars


  # Mapping  support
  library(package = "maps")
  library(package = "mapproj")
  library(package = "rgdal")
  library(package = "rgeos")

  # array -> dataframe
  library(package = "reshape2")

  library(package = "scales")

```

```{r}
# file 1

URL_root = "/maelstrom2/LOCA_GRIDDED_ENSEMBLES/LOCA_NGP/"
URL_root = "http://kyrill.ias.sdsmt.edu:8080/thredds/fileServer/LOCA_NGP/"

ref_period_file = str_c(URL_root, "climatology/DECADALS/Rdata/LOCA_NGP_DECADAL_ALL_MEMBERS_1990-1999.Rdata", sep = "")
fut_period_file = str_c(URL_root, "climatology/DECADALS/Rdata/LOCA_NGP_DECADAL_ALL_MEMBERS_2020-2029.Rdata", sep = "")

load(url(ref_period_file))
loca_hist = as_tibble(loca)
load(url(fut_period_file))
loca_futr = as_tibble(loca)
remove(loca)




loca_rcp45 = loca_futr %>% filter(str_detect(string = scenario, "4.5"))
loca_rcp85 = loca_futr %>% filter(str_detect(string = scenario, "8.5"))

loca_rcp45$Precip = loca_rcp45$Precip / 25.4
loca_rcp85$Precip = loca_rcp85$Precip / 25.4
loca_hist$Precip  = loca_hist$Precip  / 25.4


loca_rcp45$Max_Temp = loca_rcp45$Max_Temp * 9/5 + 32
loca_rcp85$Max_Temp = loca_rcp85$Max_Temp * 9/5 + 32
loca_hist$Max_Temp  = loca_hist$Max_Temp  * 9/5 + 32

loca_rcp45$Min_Temp = loca_rcp45$Min_Temp * 9/5 + 32
loca_rcp85$Min_Temp = loca_rcp85$Min_Temp * 9/5 + 32
loca_hist$Min_Temp  = loca_hist$Min_Temp  * 9/5 + 32

loca_rcp45$Avg_Temp = loca_rcp45$Avg_Temp * 9/5 + 32
loca_rcp85$Avg_Temp = loca_rcp85$Avg_Temp * 9/5 + 32
loca_hist$Avg_Temp  = loca_hist$Avg_Temp  * 9/5 + 32


loca_rcp45$Precip_Anom = loca_rcp45$Precip - loca_hist$Precip
loca_rcp85$Precip_Anom = loca_rcp85$Precip - loca_hist$Precip

loca_rcp45$Max_Temp_Anom = loca_rcp45$Max_Temp - loca_hist$Max_Temp
loca_rcp85$Max_Temp_Anom = loca_rcp85$Max_Temp - loca_hist$Max_Temp

loca_rcp45$Min_Temp_Anom = loca_rcp45$Min_Temp - loca_hist$Min_Temp
loca_rcp85$Min_Temp_Anom = loca_rcp85$Min_Temp - loca_hist$Min_Temp

loca_rcp45$Avg_Temp_Anom = loca_rcp45$Avg_Temp - loca_hist$Avg_Temp
loca_rcp85$Avg_Temp_Anom = loca_rcp85$Avg_Temp - loca_hist$Avg_Temp


lon = unique(loca_rcp45$lon)
lat = unique(loca_rcp45$lat)

loca_hist_agg = loca_hist %>% group_by(lat, lon, scenario, decade) %>% summarize(Ens_Precip        = mean(Precip,   na.rm = TRUE),
                                                               Ens_Max_Temp       = mean(Max_Temp, na.rm = TRUE),
                                                               Ens_Min_Temp       = mean(Min_Temp, na.rm = TRUE),
                                                               Ens_Avg_Temp       = mean(Avg_Temp, na.rm = TRUE),
                                                               Ens_Precip_SD      = sd(Precip,     na.rm = TRUE),
                                                               Ens_Max_Temp_SD    = sd(Max_Temp,   na.rm = TRUE),
                                                               Ens_Min_Temp_SD    = sd(Min_Temp,   na.rm = TRUE),
                                                               Ens_Avg_Temp_SD    = sd(Avg_Temp,   na.rm = TRUE))

loca_rcp45_agg = loca_rcp45 %>% group_by(lat, lon, scenario, decade) %>% summarize(Ens_Precip           = mean(Precip,        na.rm = TRUE),
                                                                 Ens_Max_Temp         = mean(Max_Temp,      na.rm = TRUE),
                                                                 Ens_Min_Temp         = mean(Min_Temp,      na.rm = TRUE),
                                                                 Ens_Avg_Temp         = mean(Avg_Temp,      na.rm = TRUE),
                                                                 Ens_Precip_SD        = sd(Precip,          na.rm = TRUE),
                                                                 Ens_Max_Temp_SD      = sd(Max_Temp,        na.rm = TRUE),
                                                                 Ens_Min_Temp_SD      = sd(Min_Temp,        na.rm = TRUE),
                                                                 Ens_Avg_Temp_SD      = sd(Avg_Temp,        na.rm = TRUE),
                                                                 Ens_Precip_Anom      = mean(Precip_Anom,   na.rm = TRUE),
                                                                 Ens_Max_Temp_Anom    = mean(Max_Temp_Anom, na.rm = TRUE),
                                                                 Ens_Min_Temp_Anom    = mean(Min_Temp_Anom, na.rm = TRUE),
                                                                 Ens_Avg_Temp_Anom    = mean(Avg_Temp_Anom, na.rm = TRUE),
                                                                 Ens_Precip_Anom_SD   = sd(Precip_Anom,     na.rm = TRUE),
                                                                 Ens_Max_Temp_Anom_SD = sd(Max_Temp_Anom,   na.rm = TRUE),
                                                                 Ens_Min_Temp_Anom_SD = sd(Min_Temp_Anom,   na.rm = TRUE),
                                                                 Ens_Avg_Temp_Anom_SD = sd(Avg_Temp_Anom,   na.rm = TRUE))


loca_rcp85_agg = loca_rcp85 %>% group_by(lat, lon, scenario, decade) %>% summarize(Ens_Precip           = mean(Precip,        na.rm = TRUE),
                                                                 Ens_Max_Temp         = mean(Max_Temp,      na.rm = TRUE),
                                                                 Ens_Min_Temp         = mean(Min_Temp,      na.rm = TRUE),
                                                                 Ens_Avg_Temp         = mean(Avg_Temp,      na.rm = TRUE),
                                                                 Ens_Precip_SD        = sd(Precip,          na.rm = TRUE),
                                                                 Ens_Max_Temp_SD      = sd(Max_Temp,        na.rm = TRUE),
                                                                 Ens_Min_Temp_SD      = sd(Min_Temp,        na.rm = TRUE),
                                                                 Ens_Avg_Temp_SD      = sd(Avg_Temp,        na.rm = TRUE),
                                                                 Ens_Precip_Anom      = mean(Precip_Anom,   na.rm = TRUE),
                                                                 Ens_Max_Temp_Anom    = mean(Max_Temp_Anom, na.rm = TRUE),
                                                                 Ens_Min_Temp_Anom    = mean(Min_Temp_Anom, na.rm = TRUE),
                                                                 Ens_Avg_Temp_Anom    = mean(Avg_Temp_Anom, na.rm = TRUE),
                                                                 Ens_Precip_Anom_SD   = sd(Precip_Anom,     na.rm = TRUE),
                                                                 Ens_Max_Temp_Anom_SD = sd(Max_Temp_Anom,   na.rm = TRUE),
                                                                 Ens_Min_Temp_Anom_SD = sd(Min_Temp_Anom,   na.rm = TRUE),
                                                                 Ens_Avg_Temp_Anom_SD = sd(Avg_Temp_Anom,   na.rm = TRUE))

loca_futr_agg = rbind(loca_rcp45_agg,loca_rcp85_agg)
```




# 5 Import Mapping Data

US and Lake Data is Available in R.  Canadian Data isn't so we have a custom polygon library

```{r}
load(file = "./canadian_provinces.Rdata")
# inported field is canada_data = canadian_provinces


usa_data   = map_data(map="state", 
                      xlim=c(-114.2812, -86.21875),  
                      ylim=c (33.96875,  52.78125) ) 
water_data = map_data(map="lakes", 
                      xlim=c(-114.2812,-86.21875),  
                      ylim=c( 33.96875, 52.78125) ) 

```





```{r}

 
  

    ggplot(data   =   loca_futr_agg)  + 
          
                       aes(x     = lon,
                           y     = lat,
                           fill  = Ens_Precip_Anom_SD) +
          

                       theme_bw() +
                       theme(strip.background = element_rect(fill=NA))+
  
                 facet_grid(cols = vars(scenario)) +

          
                       labs(title    = "LOCA Climate Ensemble Analyses ",
                            subtitle = str_c("Mean Annual Rainfall Anomalies",
                                             " ", 
                                             unique(loca_futr_agg$decade),
                                             sep = "")) +
      
                  labs(caption = str_c("Historical Base Period: ", unique(loca_hist_agg$decade), sep="")) + 

          
                       xlab(label = "Longitude") + 
          
                       ylab(label = "Latitude") +
            
             
                       coord_fixed(xlim   = c(min(lon), max(lon)),
                                   ylim   = c(min(lat), max(lat)),
                                   expand = FALSE,
                                   ratio = (max(lon)-min(lon)) / (max(lat)-min(lat))  )+
             
             
                       scale_fill_gradient2 ( name="in" ) +     
      
            
      
          
                       geom_raster(data    = subset,
                                   mapping = aes(x     = lon,
                                                 y     = lat,
                                                 fill  = Ens_Precip_Anom_SD))  +

                 

                       geom_polygon(data    = usa_data,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  + 
  
                      geom_polygon(data     = water_data,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  + 
            
                      geom_polygon(data     = canadian_provinces,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  



```


            

```{r}

 
  

    ggplot(data   =   loca_futr_agg)  + 
          
                       aes(x     = lon,
                           y     = lat,
                           fill  = Ens_Avg_Temp_Anom) +
          

                       theme_bw() +
                       theme(strip.background = element_rect(fill=NA))+
  
                 facet_grid(cols = vars(scenario)) +

          
                       labs(title    = "LOCA Climate Ensemble Analyses ",
                            subtitle = str_c("Mean Annual Average Temperature",
                                             " ", 
                                             unique(loca_futr_agg$decade),
                                             sep = "")) +
      
                  labs(caption = str_c("Historical Base Period: ", unique(loca_hist_agg$decade), sep="")) + 

          
                       xlab(label = "Longitude") + 
          
                       ylab(label = "Latitude") +
            
             
                       coord_fixed(xlim   = c(min(lon), max(lon)),
                                   ylim   = c(min(lat), max(lat)),
                                   expand = FALSE,
                                   ratio = (max(lon)-min(lon)) / (max(lat)-min(lat))  )+
             
             
                       scale_fill_gradient2 (low = "blue", mid = "white", high = "red", name="°F" ) +     
      
            
      
          
                       geom_raster(data    = subset,
                                   mapping = aes(x     = lon,
                                                 y     = lat,
                                                 fill  = Ens_Avg_Temp_Anom))  +

                 

                       geom_polygon(data    = usa_data,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  + 
  
                      geom_polygon(data     = water_data,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  + 
            
                      geom_polygon(data     = canadian_provinces,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  



```


```{r}

 
  

    ggplot(data   =   loca_futr_agg)  + 
          
                       aes(x     = lon,
                           y     = lat,
                           fill  = Ens_Min_Temp_Anom) +
          

                       theme_bw() +
                       theme(strip.background = element_rect(fill=NA))+
  
                 facet_grid(cols = vars(scenario)) +

          
                       labs(title    = "LOCA Climate Ensemble Analyses ",
                            subtitle = str_c("Mean Annual Minimum Temperature",
                                             " ", 
                                             unique(loca_futr_agg$decade),
                                             sep = "")) +
      
                  labs(caption = str_c("Historical Base Period: ", unique(loca_hist_agg$decade), sep="")) + 

          
                       xlab(label = "Longitude") + 
          
                       ylab(label = "Latitude") +
            
             
                       coord_fixed(xlim   = c(min(lon), max(lon)),
                                   ylim   = c(min(lat), max(lat)),
                                   expand = FALSE,
                                   ratio = (max(lon)-min(lon)) / (max(lat)-min(lat))  )+
             
             
                       scale_fill_gradient2 (low = "blue", mid = "white", high = "red", name="°F" ) +     
      
            
      
          
                       geom_raster(data    = subset,
                                   mapping = aes(x     = lon,
                                                 y     = lat,
                                                 fill  = Ens_Min_Temp_Anom))  +

                 

                       geom_polygon(data    = usa_data,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  + 
  
                      geom_polygon(data     = water_data,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  + 
            
                      geom_polygon(data     = canadian_provinces,
                                    mapping = aes(x     =   long, 
                                                  y     =    lat,
                                                  group =  group),
                                    fill  = NA, 
                                    color = "black")  



```

